Benchmarking neuromorphic systems with Nengo

### 2015-04-16 paper/paper.tex
Second day of writing done
I think it's a good first draft up to the results section.
Will now turn my attention to gathering results
to write about.

Benchmarking is a notoriously difficult task...

@@

@@

Go over the nengo objects

Unlike other programming models,
it's not like... for sure.
Go over symbolic dists.

Some things are suggestions.
Go over backend hints.

@@

Go over the Simulator object, and required methods.

Go over (briefly) translation of frontend to backend.
Needed: some kind of neuron model; takes current,
produces output (rates, spikes, whatever).
Needed: Ops (look at all ops)

@@

As of version 2.0.1, Nengo's test suite contains 618 tests.
Some test are unit tests though; how many?
In all, ??? can be considered ``functional'' tests,
as they construct a model,
build a simulator object (see previous sec???)
and test that the output of that model
matches the desired output,
to a certain tolerance.

@@

Nengo's test suite employs the \texttt{pytest}
testing framework \cite{???}.
??? introduce test fixtures

@@

\paragraph{\texttt{rng} \& \texttt{seed}} to manage random factors

\paragraph{\texttt{Simulator}} to switch backend

\paragraph{\texttt{nl}} to switch neuron type

\paragraph{\texttt{plt}} to save plots

\paragraph{\texttt{analytics}} to save data

\paragraph{\texttt{logger}} to save text

@@

@@

\subsubsection{Compliance}

\subsubsection{Accuracy}

\subsubsection{Speed}

@@

\paragraph{Reference}

\paragraph{Simplified reference}

\paragraph{OpenCL}

\paragraph{SpiNNaker hardware}

\paragraph{Neurogrid simulation}

>>>>

Benchmarking is a notoriously difficult task.
Benchmarks are often created by
the creators of the tools being benchmarked,
resulting in biased comparisons favoring their tool.
Benchmarking can be an inefficient process,
as the tool being benchmarked
may need changes to collect
certain performance metrics.
Even once that effort is undertaken,
benchmarks often are run a few times
and then forgotten,
quickly becoming obsolete.
We propose that the Nengo test suite
can serve as an unbiased, efficient, and future-proof
set of benchmarks for neuromorphic systems.

@@

The end result is a set of benchmarks that
are written once but run on all backends,
making new benchmarks easy to implement.
Nengo's primary goal is to
build large-scale functional models,
and contains tests of functional performance;
it is therefore not intentionally
biased in favor of any particular backend.
While there is effort involved in creating
a Nengo backend for a particular neuromorphic system,
the reason to build such a backend
is to leverage the large-scale modelling interface
provided by Nengo;
the ability to collect benchmarks
is provided essentially for free.
Finally, since the test suite exists
to ensure that Nengo continues to function
correctly, and is run frequently using
continuous integration systems,
it will continue to grow and be run frequently,
rather than becoming obsolete in the future.

@@

Nengo contains five frontend objects
that validate and store symbolic information
about the neural model that will be simulated.
A lot of reasoning can be done at this symbolic level,
which takes significant burdens off of each backend.

The primary abstraction in Nengo is the \textit{Ensemble},
which is a group of neurons (or neuron-like functions)
that collectively represent a time-varying vector of numbers.
Two parameters are mandatory: the number of neurons,
and the dimensionality of the vector being represented.
There are several optional parameters that
affect how the neurons represent the vector space;
encoders map the vector space into currents
to be injected in the neurons,
maximum firing rates can be specified for each neuron,
the type of neuron model can be specified,
and so on.

The \textit{Node} plays a similar role of representing
time-varying vectors of numbers,
but does so directly, without simulating a group of neurons.
As such, the only mandatory parameter
is what vector should be represented;
this can be specified as a constant vector,
or as a function of time.

The \textit{Connection} connects two objects
(e.g., \textit{Ensembles} or \textit{Nodes}) together.
The two objects being connected are the only mandatory parameters;
the synapse model filtering the connection
and a function applied to the vector
communicated across the connection
can optionally be specified.

The \textit{Probe} denotes that a particular
quantity in the simulation should be recorded
at a particular rate.
The object to be probed must be specified;
the attribute of that object,
a sampling rate, and a synapse model for filtering
can optionally be specified.

An important distinction between typical
parameters and those specified in Nengo
is that many are stochastic,
and not necessarily guaranteed
to be supported by every backend.
The vast majority of numeric parameters,
for example,
are most often specified as probability distributions;
an ensemble's maximum firing rate defaults to
a uniform distribution between 200 and 400 Hz,
but this could be set to a Gaussian distribution
or a discrete set of options and associated
probabilities.
Other parameters,
most notably an ensemble's neuron model,
may not be supported by a particular backend,
which may raise an error.
Even if a particular neuron model is supported,
it may be implemented differently by each backend;
Nengo does its best to approximate
the high-level specification regardless of
the neuron model that is actually used.

Finally, the \textit{Network} is a
container for the other four objects.
The only parameter on the network
is an optional integer seed;
setting this should make
all of the random factors
in the model deterministic.
Aside from managing the random seed,
the network is responsible for maintaining
a network-specific set of default parameters
for the four other objects.
This results in much shorter and less error-prone
model creation scripts.

An example script showing how Nengo can
express a functional network
with ??? neurons and ??? connections
(representing ??? connection weights)
in ??? lines of code can be seen
in Figure~\ref{???}.

@@

The role of the backend is to take
a single network, which contains
ensembles, nodes, connections, and probes,
and construct the underlying objects necessary
to implement the model specified by that network.
That implementation is exposed to Python
through a \textit{Simulator} object,
which has three required methods
and one required attribute.

The first method is \texttt{\_\_init\_\_},
which is a special Python method for initializing objects.
This method must accept a network as its first argument.
It can then accept optional arguments depending on
the capabilities and requirements of the backend;
the reference backend, for example,
accepts \texttt{dt},
which is the length of each timestep.
The purpose of \texttt{\_\_init\_\_} is to
set up the low-level system
that implements the high-level objects
contained in the provided network.
When parts of the model must be
randomly sampled, the random factors
should be identical when a network
In the reference backend,
this involves sampling from the distributions
in all of the parameters of all objects,
solving for decoding weights and connection weights,
setting up data structures for probed data,
and so on.

The second method is \texttt{run},
which advances the simulation
by the number of seconds
passed in as a required argument.
Whatever low-level structure was created
in \texttt{\_\_init\_\_} advances forward
for that many real or simulated seconds,
and importantly, any probed data is sampled
according to its sample rate.
In the reference backend,
this advances the simulation
by \texttt{time / dt} timesteps,
where \texttt{dt} is fixed at the start of the simulation
and cannot change;
other backends may have variable-length timesteps,
or may not explicitly track timesteps.

The third method is \texttt{trange},
which returns a sequence of times that correspond
to the times at which data was probed.
In the reference backend, this will be
a straightforward sequence of increasing
multiples of \texttt{dt};
other backends may simulate at variable rates.

Finally, the simulator object exposed by each backend
must have an attribute called \texttt{data},
which exposes a dictionary-like interface
to the data being probed over the course
of the simulation.
The data dictionary should,
at a minimum, provide a mapping from
probe instances to the probed data;
in other words, probe instances are keys
and the probed data are values.

These three methods and one attribute
make up the public interface
that Nengo tests expect from backends;
it is an example
of the adapter pattern \cite{???}. % Maybe lose?
While the existence of the methods
and attribute are necessary
for every Nengo model,
they are not sufficient;
backends are not required
to implement all of the neuron models
or advanced features that
are available in all other backends.
Backends are responsible for informing
the user when model cannot be implemented
on that backend.
The compliance metric (see section ???)
explicitly tracks the details of what
types of models can be implemented
by each backend.
The details of how each backend
implements high-level frontend model
are out of the scope of this paper.

@@

As of version 2.0.1, Nengo's test suite contains 618 tests.
Some test are unit tests though; how many?
In all, ??? can be considered ``functional'' tests,
as they construct a model,
build a simulator object (see previous sec???),
run that object,
and test that the output of that model
(as exposed through \texttt{trange}
and \texttt{data})
matches the desired output,
to a certain tolerance.

@@

Nengo's test suite employs the \texttt{pytest}
testing framework \cite{???}.
\texttt{pytest} enables
expressive testing of pluggable components
(such as backends)
through what are called ``test fixtures.''
Fixtures are exposed to the test suite
as arguments that can be provided
to any test function.
\texttt{pytest} inspects the function signature,
and passes an appropriate value
to the function.
This allows for boilerplate code
to be specified once,
and run for any test that might use it.
The example script in Figure~\ref{???}
demonstrates how these fixtures
look in test scripts.

@@

Nengo's test suite defines several test fixtures
to enable fuzzy testing
that collects and visualizes performance metrics
on multiple backends.

\paragraph{\texttt{rng} \& \texttt{seed}}
enable deterministic testing despite
all of the random factors that affect a model.
Test functions that generate random values
(as inputs to a model, for example)
generate those values using \texttt{rng},
which is a random number generator
with an interface identical to
NumPy's random module.
When a test function creates a network,
it passes in \texttt{seed} to make
the random components
of the frontend-to-backend translation deterministic.
Importantly, these random components
are seeded by hashing a string uniquely
identifying each test function,
ensuring that tolerances must be set
wide enough that changing the seed
should only rarely result in test failures.
The hash function is also seeded
with a global integer,
which is changed periodically to ensure
that tests have good tolerances.

\paragraph{\texttt{Simulator}}
allows for tests to be run with multiple backends.
As detailed in section ???,
each backend must expose a simulator class
that accepts a network.
Every test uses this test fixture rather than
an actual class so that
the backend-specific simulator class is used
when the test is run.
Each backend's test suite loads
the tests defined in Nengo,
but replaces the implementation of
this fixture with their own fixture
returning that particular backend's simulator class.

\paragraph{\texttt{plt}, \texttt{logger}, \& \texttt{analytics}}
allow tests to save artifacts from test runs.
These artifacts are how accuracy and speed metrics
are collected, and how the figures
in this paper are generated.
The \texttt{plt} fixture exposes the Matplotlib \cite{???}
\texttt{pyplot} interface to a test function.
These functions can then analyze and plot
the data generated in a simulation
(exposed from a simulator's \texttt{data} attribute)
to visualize the activity in a network.
Figures are saved in a specified or backend-specific directory.
The \texttt{logger} fixture exposes a logging interface
to save arbitrary text to a specified
or backend-specific directory;
these are useful for summary statistics
that can inspected manually.
For large amounts of data that
requires more analysis,
\texttt{analytics} fixture exposes an interface
to save arbitrary data (in the form of NumPy arrays)
to specified or backend-specific directories.

\paragraph{\texttt{analytics\_data}}
enables comparative testing between
two runs of the same test.
The two runs might represent two versions
of the same backend---one before a
speedup and one after---or two
entirely different backends.
The \texttt{analytics\_data} fixture
provides the results from those two
test runs to a single test function,
allowing for comparisons
between the two results saved
with the \texttt{analytics} fixture.
See Figure~\ref{???} for a
concrete example of this.

@@

Benchmarking involves collecting performance metrics
for two or more comparable systems.
Currently, the Nengo test suite collects
three types of performance metrics
across five backends.

@@

\paragraph{Compliance\textcolon}

\paragraph{Accuracy\textcolon}

\paragraph{Speed\textcolon}

@@

\paragraph{Reference\textcolon}

\paragraph{Simplified reference\textcolon}

\paragraph{OpenCL\textcolon}

\paragraph{SpiNNaker hardware\textcolon}

\paragraph{Neurogrid simulation\textcolon}
